### 2019-08-30 paper/boneyard.tex
finished neuroimaging methods

\subsection*{Evaluation metrics}
We evaluate our approach to extracting dynamic correlations and
higher-order correlations using several metrics detailed next.  First,
we generated synthetic data using known time-varying correlations, and
then we evaluated the fidelity with which Equation~\ref{eqn:timecorr}
could recover those correlations (for synthetic datasets with
different properties, and using different kernels to define the
weights; Fig.~\ref{fig:kernels}).  We then turned to a series of
analyses on a (real) neuroimaging dataset where the ground truth
correlations were \textit{not} known.  We evaluated whether the
recovered correlations could be used to accurately label held-out
neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
decoding) as a proxy for
gauging how much explanatory power the recovered correlations held
with respect to the observed data.

\subsubsection*{Generating synthetic data}
To explore recovery of a constant covariance (Fig.~\ref{fig:synthetic_data},  a.), we generated syntheic
data sampled from a constant covariance matrix. To do this, we created one random
covariance matrix, $K$, with 50 features, and for each of the 300 timepoints  we
sampled from a Gaussian distribution centered on $K$.  Similarly, we generated synthetic data
sampled from a random covariance matrix (Fig.~\ref{fig:synthetic_data}, b.) by creating a new random
covariance matrix $K(t)$, for each of the 300 timepoints and sampled from a
Gaussian distribution centered on $K(t)$.

To generate synthetic data from a dynamically changing covariance
matrix (Fig.~\ref{fig:synthetic_data},  c.),
we generated two random covariance
matrices, $K_{1}$ and $K_{2}$.  We then computed a weighted average covariance matrix
for each of the 300 timepoint, $K(t)$, 
by taking the linearly spaced weights ($w$) of the two random matrices, 
\begin{align}
K(t) = w(t) * K_{1} + (1 - w(t)) *  K_{2}, \\\label{eqn:ramping}
\end{align}
and for each of the 300 timepoints sampled from a
Gaussian distribution centered on $K(t)$.

Lastly, for the synthetic data containing block structure (Fig.~\ref{fig:synthetic_data},  d.), we followed the same
process of creating synthetic data sampled from a constant covariance
matrix (see above) but sampled from a new random covariance matrix
after 60 consecutive timepoints.  We then pieced the blocks together
to create a sythetic dataset with 300 total timepoints but drawn from
5 separate covariance matrices.

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}

We applied timecorr, using delta and gaussian (width = 10) kernels
Fig.~\ref{fig:kernels}) to each of these 
synthetic datasets, then correlated each recovered
correlation matrix with the ground truth.  We repeated this process 10
times and explored how recovery varies
with the kernel and the specific structure of the data. For the
ramping synthetic dataset (Fig.~\ref{fig:synthetic_data},  c.)  and for the
block synthetic dataset (Fig.~\ref{fig:synthetic_data},  d.)  we made further
comparisons of the timecorr recovered correlation matrices. We
compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
c.) from Equation~\ref{eqn:ramping}. We also compared the block recovered correlation matrices in to
the block specific covariance matrix (Block 1-5,
Fig.~\ref{fig:synthetic_data},  d.).

A major challenge to studying such patterns is that typically neither
the correlations nor the hierarchical organizations of those
correlations may be directly observed.  Rather, these fundamental
properties must be inferred indirectly by examining the observable
parts of the system-- e.g., the behaviors of the individual units of
that system.  Here we propose a series of mathematical operations that may be used to approximate dynamic correlations at a range of scales (i.e., orders of interaction).

There are two basic steps to our approach (Fig.~\ref{fig:methods_fig}).  In the first step, we take
a number-of-timepoints ($T$) by number-of-features ($F$)
\textit{matrix of observations} ($\mathbf{X}$) and we return a $T$ by
$\frac{F^2 - F}{2}$ \textit{matrix of dynamic correlations}
($\mathbf{Y}$).  Here $\mathbf{Y_0}$ describes, at each moment, how
all of the features (columns of $\mathbf{X}$) are inferred to be
interacting.  (Since the interactions are assumed to be non-recurrent
and symmetric, only the upper triangle of the full correlation matrix
is computed.)  In the second step, we project $\mathbf{Y_0}$ onto an
$F$-dimensional space, resulting in a new $T$ by $F$ matrix
$\mathbf{Y_1}$.  Note that $\mathbf{Y_1}$ contains information about
the correlation dynamics present in $\mathbf{X}$, but represented in a
compressed number of dimensions.  By repeatedly applying these two
steps in sequence, we can examine and explore higher order dynamic
correlations in $\mathbf{X}$.

To understand the neuronal computations that support cognition, we
must understand the cooperative dynamics of populations of neurons.
These populations of neurons interact within each brain structure, and
the structures interact to form complex and dynamic networks. These
interactions, at each scale, vary according to the functions our
brains are carrying out and recent work has shown these dynamic
complex patterns support consciousness~\citep{DemeEtal19}.

In the last several decades, advances in functional magnetic resonance
image (fMRI) analyses have evolved to better characterize these
interactions of brain structures.  As~\cite{Turk13} outlines,
analyses using multivariate patterns of activity have given an advance
over univariate activity patterns because they allow relative
contributions of voxels to be combined and better assess distrubted
representations. Resting-state connectivity (RS) fMRI, which is the
temporal correlation of regions during rest, has shed light on the
rich and complex spatiotemporal organization of spontaneous brain
activity.  Functional connectivity (FC) has further characterized the
cognitive state dependency of this network organization.

Recent work has shown that FC fluctuates over time~\citep{ChanGlov10}~\citep[for review]{LuriEtal18}, and the assumed stationarity of these analyses may be too
simple to capture the dynamic nature of brain activity.  Additionally,
recent work has shown that temporal
variability in functional connectivity predicts attention task
performance~\citep{FongEtal19} and that
dynamic correlations between multivariate voxel patterns can add an
additional boost compared to static multivariate voxel patterns alone~\citep{MannEtal18}.

  Most complex systems reflect dynamic interactions between myriad
  evolving components (e.g., interacting molecules, interacting brain
  systems, interacting individuals within a social network or
  ecological system, coordinated components within a mechanical or
  digital device, etc.).  Despite that these interactions are central
  to the full system's behavior (e.g., removing a component from the
  full system can change the entire system's behavior), dynamic
  interactions cannot typically be directly measured.  Rather, the
  interactions must be inferred through their hypothesized role in
  guiding the dynamics of system components.  Here we use a
  model-based approach to inferring dynamic interactions from
  timeseries data.  In addition to examining first-order interactions
  (e.g., between pairs of components) we also examine higher-order
  interactions (e.g., that characterize mirrored structure in the
  patterns of interaction dynamics displayed by different subsets of
  components).  We apply our approach to two datasets.  First, we use
  a synthetic dataset, for which the underlying dynamic interactions
  are known, to show that our model recovers those ground-truth
  dynamic interactions.  We also apply our model to a neuroimaging
  dataset and show that the high-order dynamic interactions exhibited
  by brain data vary meaningfully as a function of the cognitive
  ``richness'' of the stimulus people are experiencing.

%should we introduce the name levelup in the introduction? ie. "in the second step, levelup, we use..." -ecw
%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%

The mathematical theories behind the timecorr approach for calculating the moment-by-moment dynamic correlations in systems are described here. The function takes a number-of-timepoint by number-of-features matrix as well as a weight function to determine a moment-by-moment correlations matrix. To calculate the correlations, expectation values are found using equation 1:

\begin{equation}
\mathrm{corr}_{x,y}(t) = \frac{1}{T} \sum_{t} \frac{\left(x(t)-\mathbb{E}[x]\right)\left(y(t)-\mathbb{E}[y]\right)}{\sigma_{x}(t)\sigma_{y}(t)}
\end{equation}

%should the variables in this equation/description change to match the description in the intro?
Here, x is the timepoint component of the matrix and y is the feature component. Timepoint by timepoint, x by x, at each given moment (t) in the matrix is subtracted from the expectation value of the full timepoint matrix. The same process occurs for the features points (y) at each timepoint (t). The product of these are then divided by the standard deviation at that moment for both the timepoint and the feature. Only the upper triangle of the matrix is calculated as the interactions are assumed to be both symmetric and non-recurrent.

Two weight parameters have been used to calculate the connectivity of the data are a Gaussian weight parameter, equation 2, and a LaPlace weights parameter, equation 3.

%label as Gaussian? 
\begin{equation}
\mathcal{N}\left( x | \mu, \sigma \right) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left(-\frac{|| x - \mu ||^2}{2\sigma^2}\right)
\end{equation}
where sigma squared is the variance and mu is the expected value.

%label as LaPlace? 
\begin{equation}
\mathcal{L}\left( x | \mu, \lambda \right) = \frac{1}{2 \lambda}\left( x | \mu, \lambda \right) = \exp\left(-\frac{|x - \mu|}{\lambda}\right)
\end{equation}
where mu is the expectation values, lambda is a defined decay constant.

%figure showing the difference b/w two weights in same data set? Using the synthetic data?

When using the Gaussian weights parameter timepoints are more closely correlated to timepoints further away in the matrix than when using then the LaPlace weight function.

%Plot of kernels as Fig 1?

% New section? Mathematical Methods of Levelup? 
In order to understand how the data is interconnected we use the function levelup. Levelup uses the same input as the timecorr function, a timepoint-by-function matrix.

%Diagram of input-->timecorr-->PCaA= levelup?

In the levelup method the the data first goes through the timecorr process and then is reduced using Principal Component Analysis (PCA). By utilizing PCA the output data maintains the same features as the input method. Not only does this allow for a sanity check, but also it allows for repeated use of the function. Using PCA or other reduction techniques is needed as without it the output data would get exponentially large. On the flip side using PCA, or any dimensionality reduction technique, makes the output data less accurate as information is lost.

%PCA equation?
%figure/ chart of how many levels we "go up" before data gets too muddled- depends on the correlation?

Timepoints that share high similarity across all subjects have a high probability of being stimulus driven.

To compute the correlations of multiple subjects the across mode of both timecorr and levelup are used. The across function compares one voxel in a sample to the averages of that voxel in the rest of the dataframe. This occurs throughout every voxel in the dataset. The output of the across mode is a correlation matrix that is a number of timepoints by features matrix.

%F^2-F/2 +F